---
title: "Coronavirus in South Africa"
output:
  html_document:
    df_print: paged
---

This is a visual that looks at comparing the spread of the coronavirus in countries outside of China.  

```{r echo=FALSE, include=FALSE}
# packages to use

library(tidyverse)
library(hrbrthemes)
library(ggtext)
```

The file used here is soucred from this [CSSE at Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv) GitHub repo. Date stamp for this file is 27 March 2020 9:03:24 AM.

### Loading the file ###

```{r}

# reading in the csv file
coronavirus_source <- read.csv('time_series_covid19_confirmed_global_2020_03_27.csv')

# unpivot the date columns into a single column called days
coronavirus <- coronavirus_source %>% gather(days,"counts", -c(Province.State, Country.Region, Lat, Long)) %>% filter(!counts == 0)

# remove the X character on the dates
coronavirus$days <- str_replace(coronavirus$days, "X", "")

# to see the number of data logged for South Africa
coronavirus %>% filter(Country.Region == 'South Africa')

# since South Africa is on day 22 since the virus was first reported in the country
coronavirus_22 <- coronavirus %>% group_by(Country.Region) %>% slice(1:22)
```

Now I need to check the numbers from Italy from day 1 to day 22. In Italy, at day 22, they had 20 reported cases of the Coronavirus.

```{r}
coronavirus_22 %>% filter(Country.Region == 'Italy') 

coronavirus_latest_22 <-  coronavirus_22 %>% arrange(desc(counts)) %>% slice(1) %>% arrange(desc(counts)) %>% filter(!Country.Region == 'China', counts >= 200) %>% group_by(Country.Region) %>% summarise(totalcases = sum(counts))
```

### Create the visual ###

Creating the visual to see how South Africa compares to the countries outside of China, at day 22 of the Coronavirus since the first case(s) were reported. This is a ggplot2 visual.

```{r error=FALSE, warning=FALSE, fig.width=10, fig.height=8.5}
coronavirus_latest_22 %>% 
  ggplot(aes(x = reorder(Country.Region, totalcases), y = totalcases)) +
  geom_col(color = "black") +
  geom_col(data = coronavirus_latest_22 %>% filter(Country.Region == "South Africa"), fill = "red", color = "black") +
  geom_text(data = coronavirus_latest_22, aes(label=totalcases), hjust=-0.2, nudge_y= 1, size = 4) +
  coord_flip() +
  theme_ipsum(grid = F) +
  labs(x = "Countries", y = "Total Number of Cases After 22 Days", 
       title = "**Total Number of <i style='color:#3F704D'>*Coronavirus*</i> Cases 22 Days After the <i style='color:red'>*First Reported Case(s)*</i>**  
    ", 
       subtitle = "Looking at countries outside of China that had 200 or more reported cases, 22 days after the 1st case(s).",
       caption = "Data Source: https://data.humdata.org/dataset/novel-coronavirus-2019-ncov-cases\nBy S'busiso Mkhondwane") + 
  theme(text = element_text(family = "Roboto Condensed"),
        plot.title = element_markdown(lineheight = 1.2, colour = "black", face = "bold", hjust = 0.5, size = 15), 
        plot.subtitle = element_markdown(hjust = 0.5, face = "bold"),
        panel.grid.minor.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none"
  ) 
 
#ggsave(filename = here::here("data/Coronavirus_after_22_days.png"),
#       height = 8, width = 14)
```

